Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-03-27 manuscript.tex
Finished minor lab edits

Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior. Supervised recordings from past work (Airaksinen et al., 2022, 2020; Franchak et al., 2021), whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors. Prior work focused on ``free play'', in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move. However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position. Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time? Assessing the validity of temporally \emph{distal} periods is a crucial step to establish whether automatic classification can be used to measure body position across a day.

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the hip and a pocket just above the ankle were sewn on the outside of the right and left legs of the garment. Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest.

@@

Figure \ref{fig:exemplar-timeline}) shows an exemplar timeline of the entire procedure and recording periods for a single participant. On the day of the visit, a researcher arrived at the participant's home in the morning between and prepared all the equipment at the doorstep. In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory). To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in Franchak et al. (2021). All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door. The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for ground truth human annotation of body position. At the start of this ``guided call'', the caregiver was instructed to place the camera in an area that captured the majority of the room. Next, they were asked to put the pair of leggings and shirt on their infant, with the researcher providing guidance about how to correctly orient the garments.

@@

Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes. The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline})) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline})) so that those times could be excluded from analysis. The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.

@@

Supine was coded when the infant was lying on their back or was reclined up to a 45 degree angle. Supine was also coded in the rare cases when the infant was laying on their side. Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling. We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine). Upright was coded when the infant was standing on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support). When an infant was carried by a caregiver, held was coded. However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting). Times during the video when the infant was out of view were excluded. Periods when the leggings were being adjusted or taken off the infant were also excluded, as were transitions between body positions.

@@

Datasets containing one row per 4-s window were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package (Liaw et al., 2002). Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The \emph{tidymodels} set of packages in R (Kuhn \& Wickham, 2020) was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the \emph{predict} function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

@@

\caption{Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period. Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models. Each blue circle represents that accuracy for each recording session. Horizontal black bars indicates the mean across sessions.}\label{fig:metrics}
\end{figure}

@@

The first set of analyses use data from the proximal comparison to determine the ``best case'' accuracy of the models, training and testing on similar types of data. The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work (Franchak et al., 2021), we compared two types of models: \emph{group models} and \emph{individual models}. To assess the accuracy of body position classification for each recording session, we reserved the last 25\% of a session's proximal comparison data as the testing set. The testing set was never used as training data. The two types of models differed based on what data were used during training, but the same random forest modeling procedure was used in both cases. A group model was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75\% of a recording session, then tested against the last 25\% of the same session. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \ref{fig:metrics}, Table \ref{tab:metrics-table}, and Table \ref{tab:metrics-by-class} summarize the performance of group and individual models using standard metrics for classification.

@@

Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92). Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies (Airaksinen et al., 2020; Franchak et al., 2021), both models approached the level of agreement found between two human coders (\emph{M} = .906). Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model. Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy. Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93. Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was considerably lower than in individual models (minimum = 0.73). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

@@

Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with group kappas (\emph{M} = 0.75) somewhat worse compared with individual kappas (\emph{M} = 0.82). Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

@@

Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another. Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?). High PPV means that we can be confident in the event label. Table \ref{tab:metrics-by-class} shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. All values were sufficient (\textgreater{} .8). Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types. Table \ref{tab:metrics-table} shows the overall sensitivity and PPV across classes.

@@

In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period (Figure \ref{fig:exemplar-timeline}). A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to \textasciitilde{} 5 s in either direction). This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work (Franchak et al., 2021; Yao et al., 2019). For each session, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding. Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.

Whereas the proximal analyses used all 34 sessions, this was not possible in the distal comparison. Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period. Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording. Three additional sessions were excluded because a caregiver accidentally turned off the video camera (\emph{n} = 1) or (purposefully) left the house (\emph{n} = 2). This left 22 sessions with usable distal comparison data.

@@

Visual inspection of Figure \ref{fig:part2overall} indicated two extreme outliers, which we marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions. The ``gray square'' outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording. In contrast, the ``gray diamond outlier'' had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as ``upright'' but the models predicted it as ``held''. Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infants moved while walking upright.

@@

\caption{Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison period. Each point shows the mean and SE for a single participant's recording session for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The two outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
\end{figure}

@@

Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification. Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated the agreement analysis (correlating human-coded body position prevalence with model-predicted prevalence) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}). Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera. Bins were included only if there was \textgreater{} 7 minutes of usable data.

@@

\caption{Full-day timelines for each individual recording session, split by (A) younger infants and (B) older infants. Each participant's timeline shows a stacked bar graph with the proportion of time spent in each of the body position for every 5-minute period throughout the day, based on group model predictions of body position. The x-axis shows time of day. Caregiver-reported naps are marked by gray bars; blank gaps indicate caregiver-reported times that the sensor garment was removed for diaper changes, baths, or trips out of the house.}\label{fig:timelines}
\end{figure}

@@

Figure \ref{fig:timelines} depicts body position timelines across the day for each session, divided into younger (A) and older infants (B). Predictions from the group model were used to ensure that motion data were classified consistently across all sessions. Session start times ranged from 09:15 to 13:20 with a median of 10:25. Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group). With two exceptions, recordings lasted until the infants' bed time. Older participant \#1 had the equipment picked up on the same day rather than the next day, so the recording ended at 17:00. Older participant \#11 wore the equipment in the morning, but the family left the house at 10:15 and remained out for the rest of the day, choosing not to put the equipment back on when they returned home in the late afternoon. Among those participants who wore the equipment until bedtime, recordings ended between 17:25 and 22:15 with a median of 19:20.

@@

Considering that body position annotation takes 2-5 hours of labor per 1 hour of behavior, the 207 hours of data we recorded would have taken 413-1033 hours of labor to annotate from video. However, only an estimated 64-160 hours of labor was needed to annotate the initial hour of each session's video that was used to train the machine learning models---an immense savings in the human cost of annotation that allows the method to scale to larger sample sizes and recording durations.

@@

In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age. Past work using sparse EMA surveys across the day (Franchak, 2019) found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright. Prone time changed little. Table \ref{tab:agetable} summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.

The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically. In Franchak (2019), upright time was 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months. Results from the current study were similar: younger infants' sitting time estimated from full-day group models was 5.41\% compared with 18.24\% for the older group. In Franchak (2019), prone time was 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months. In the current study, prone time was 14.03\% for younger infants and 15.09\% for older infants. Held, sitting, and supine are more difficult to compare because prior work (Franchak, 2019) considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively. Further complicating the comparison, Franchak (2019) included a reclined category that we would classify as either supine or sitting (whichever it was closer to). Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \ref{tab:agetable}). Sitting was greater for older infants (44.39\%) compared with younger infants (28.91\%). Moreover, sitting was the most frequent body position followed by upright in 12-month-olds (Franchak, 2019), just as we observed in the current study. Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.

@@

Across three analyses, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors. Both group and individual models trained during the proximal comparison period performed well on testing data collected from the same recording period. Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes. Examining full recordings showed that the new method allows us to capture more data (median = 6.14 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it. Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.

@@

Counterintuitively, the two outlier participants increased our confidence that the body position classification method works as intended. At first glance, seeing low agreement rates for two sessions would suggest that the models perform poorly at predicting some \emph{infants'} behavior. Instead, we found that for those infants, errors were restricted to two positions during a single, long-lasting event, while the other three body positions continued to be classified correctly. In other words, the models failed to predict a particular \emph{event} for each of the two infants. Unfortunately, those events lasted a long time leading to poor overall agreement. In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficult of categorizing behaviors that do not always fit neatly into discrete classes (e.g., continuous body position angles categorized as either supine or sitting).

>>>>

Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior. Supervised recordings from past work (Airaksinen et al., 2022, 2020; Franchak et al., 2021), whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors. Prior work focused on ``free play'', in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move. However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position. Indeed different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to be counted as equal. Moreover, models must be able to new variants of behaviors that occur over the course of natural behavior that did not appear in the training data. Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time? Assessing the validity of temporally \emph{distal} periods is a crucial step to establish whether automatic classification can be used to measure body position across a day.

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest.

@@

Figure \ref{fig:exemplar-timeline}) shows an exemplar timeline of the entire procedure and recording periods for a single participant. On the day of the visit, a researcher arrived at the participant's home in the morning and prepared all the equipment at the doorstep. In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory). To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in Franchak et al. (2021). All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door. The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for ground truth human annotation of body position. At the start of this ``guided call'', the caregiver was instructed to place the camera in an area that captured the majority of the room. Next, they were asked to put the pair of leggings and shirt on their infant, with the researcher providing guidance about how to correctly orient the garments.

@@

Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, diaper changes, and trips out of the house. The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline}) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline}) so that those times could be excluded from analysis. The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.

@@

Supine was coded when the infant was lying on their back or was reclined up to a 45 degree angle. Supine was also coded in the rare cases when the infant was laying on their side. Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling. We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine). Upright was coded when the infant was standing or squatting on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support). When an infant was carried by a caregiver, held was coded. However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting). Times during the video when the infant was out of view were excluded. Periods when the leggings were being adjusted or taken off the infant were also excluded, as were transitions between body positions.

@@

The resulting windowed datasets were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package (Liaw et al., 2002). Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The \emph{tidymodels} set of packages in R (Kuhn \& Wickham, 2020) was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550; however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the \emph{predict} function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

@@

\caption{Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period. Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models. Each blue circle represents that accuracy for each recording session. Horizontal black bars indicate the mean across sessions.}\label{fig:metrics}
\end{figure}

@@

The first set of analyses use data from the proximal comparison to determine the ``best case'' accuracy of the models, training and testing on similar types of data. The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work (Franchak et al., 2021), we compared two types of models: \emph{group models} and \emph{individual models}. To assess the accuracy of body position classification for each recording session, we reserved the last 25\% of a session's proximal comparison data as the testing set. The testing set was never used as training data. The two types of models differed based on what data were used during training, but the same random forest modeling procedure was used in both cases. A group model was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75\% of a recording session, then tested against the (temporally) last 25\% of the same session. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \ref{fig:metrics}, Table \ref{tab:metrics-table}, and Table \ref{tab:metrics-by-class} summarize the performance of group and individual models using standard metrics for classification.

@@

Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92). Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies (Airaksinen et al., 2020; Franchak et al., 2021), both models approached the level of agreement found between two human coders (\emph{M} = .906). Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model. Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy. Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93). Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was considerably lower than in individual models (minimum = 0.73). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

@@

Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with group kappas (\emph{M} = 0.75) somewhat worse compared with individual kappas (\emph{M} = 0.82). Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

@@

Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another. Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded upright events?). High PPV means that we can be confident in the event label. Table \ref{tab:metrics-by-class} shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types. Table \ref{tab:metrics-table} shows the overall sensitivity and PPV across classes.

@@

In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period. A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to \textasciitilde{} 5 s in either direction). This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work (Franchak et al., 2021; Yao et al., 2019). For each session, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding. Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.

Whereas the proximal analyses used all 34 sessions, this was not possible in the distal comparison. Because the start of the visit was scheduled during a time when the infant was awake, it was common for a nap to be needed following the proximal period. Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording. Three additional sessions were excluded because a caregiver accidentally turned off the video camera (\emph{n} = 1) or (purposefully) left the house (\emph{n} = 2). This left 22 sessions with usable distal comparison data.

@@

Visual inspection of Figure \ref{fig:part2overall} indicated two extreme outliers, which we marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions. The ``gray square'' outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording. In contrast, the ``gray diamond outlier'' had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as ``upright'' but the models predicted it as ``held''. Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infants moved while walking upright.

@@

\caption{Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison period. Each point shows the mean and SE for a single recording session for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The two outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
\end{figure}

@@

Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification. Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated the agreement analysis (correlating human-coded body position prevalence with model-predicted prevalence) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}). Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera. Bins were included only if there was \textgreater{} 7 minutes of usable data.

@@

\caption{Full-day timelines for each individual recording session, split by (A) younger infants and (B) older infants. Each participant's timeline shows a stacked bar graph with the proportion of time spent in each of the body positions for every 5-minute period throughout the day, based on group model predictions of body position. The x-axis shows time of day. Caregiver-reported naps are marked by gray bars; blank gaps indicate caregiver-reported times that the sensor garment was removed for diaper changes, baths, or trips out of the house.}\label{fig:timelines}
\end{figure}

@@

Figure \ref{fig:timelines} depicts body position timelines across the day for each session, divided into younger (A) and older infants (B). Predictions from the group model were used to ensure that motion data were classified consistently across all sessions. Moreover, for infants who did not display all 5 body positions in the training period, group models were necessary to be able to predict those behaviors across the entire day. Session start times ranged from 09:15 to 13:20 with a median of 10:25. Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group). With two exceptions, recordings lasted until the infants' bedtime. Older participant \#1 had the equipment picked up on the same day rather than the next day, so the recording ended at 17:00. Older participant \#11 wore the equipment in the morning, but the family left the house at 10:15 and remained out for the rest of the day, choosing not to put the equipment back on when they returned home in the late afternoon. Among those participants who wore the equipment until bedtime, recordings ended between 17:25 and 22:15 with a median of 19:20.

@@

Considering that body position annotation takes 2-5 hours of labor per 1 hour of behavior, the 207 hours of data we recorded would have taken 413-1033 hours of labor to annotate from video. However, only an estimated 64-160 hours of labor was needed to annotate the initial hour of each session's video that was used to train the machine learning models---an immense savings in the human labor cost of annotation that allows the method to scale to larger sample sizes and recording durations.

@@

In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work (Franchak, 2019). Table \ref{tab:agetable} summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models. Past work using sparse EMA surveys across the day (Franchak, 2019) found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright. Prone time changed little. The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically. In Franchak (2019), upright time was 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months. Results from the current study were similar: younger infants' upright time estimated from full-day group models was 5.41\% compared with 18.24\% for the older group. In Franchak (2019), prone time was 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months. In the current study, prone time was 14.03\% for younger infants and 15.09\% for older infants. Held, sitting, and supine are more difficult to compare because prior work (Franchak, 2019) considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively. Further complicating the comparison, Franchak (2019) included a reclined category that we would classify as either supine or sitting (whichever it was closer to). Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \ref{tab:agetable}). Sitting was greater for older infants (44.39\%) compared with younger infants (28.91\%). Moreover, sitting was the most frequent body position followed by upright in 12-month-olds (Franchak, 2019), just as we observed in the current study. Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.

@@

Across three analyses, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors. Both group and individual models trained during the proximal comparison period performed well on testing data collected from the same recording period. Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes. Examining full recordings showed that the new method allows us to capture more data (median = 6.14 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it. Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for capturing individual differences and describing developmental changes in motor behavior.

@@

Counterintuitively, the two outlier participants increased our confidence that the body position classification method works as intended. At first glance, seeing low agreement rates for two sessions would suggest that the models perform poorly at predicting some \emph{infants'} behavior. Instead, we found that for those infants, errors were restricted to two positions during a single, long-lasting event, while the other three body positions continued to be classified correctly. In other words, the models failed to predict a particular \emph{event} for each of the two infants. Unfortunately, those events lasted a long time leading to poor overall agreement. In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficulty of categorizing behaviors that do not always fit neatly into discrete classes (e.g., continuous body position angles categorized as either supine or sitting).
